Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-16 paper/current/main.tex
adjust math

\begin{equation}
	\label{eq:gravity_basic}
	\hat{T_{ij}} = Cm_{i}m_{j}f(r_{ij})
\end{equation}

Where the \textit{expected flux}, $\hat{T_{ij}}$, or expected number of authors who were affiliated with both organizations during the time period is proportional to the organization's sizes (the sum of both mobile and non-mobile authors, see Data and Methods), $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$.
The distance function is defined in terms of either the geographic distance between organizations or the embedding distance. 
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function, selected as the best performing for each case (Fig.~\ref{fig:supp:predict_emb_bymodel} and Fig.~\ref{fig:supp:predict_geo_bymodel}).
We argue that the performance of a distance measure can be observed from its ability to explain expected mobility between organizations and to make predictions using the gravity model.

@@

\begin{equation}
	\label{eq:gravity_basic}
	\hat{T_{ij}} = Cm_{i}m_{j}f(r_{ij})
\end{equation}

where $m_{i}$ and $m_{j}$ are the organization masses, or the yearly-average count of unique authors affiliated with each organization during the time window, and $f(r_{ij})$ is a distance-decaying function  between the two organizations. There are two popular forms of the $f(r_{ij})$. The first one is power-law function with $f(r_{ij})= r^{-\alpha}_{ij} \ (\alpha > 0)$, and the another one is exponential function with $f(r_{ij}) = e^{-\beta r_{ij}} \ (\beta > 0)$ \autocite{chen2015distance}.  $f(r_{ij})$, $C$ can be chosen as to fit the data with the simple regression.

We consider separate variants of $f(r_{ij})$ for geographic distance, $g_{ij}$, and the embedding distance, $d_{ij}$, report the best-fit model of each distance.
For geographic distance, we use the power-law function of the gravity law, with $f(g_{ij})= g^{-\alpha}_{ij}$, where $\alpha$ is estimated from the data.
For embedding distance, we use the exponential function, with $f(d_{ij}) = e^{-\beta d_{ij}}$, where $\beta$ is estimated from the data.
To estimate parameters $C$, $\alpha$ and $\beta$ are estimated by log-transforming each variant of \ref{eq:gravity_basic}, obtaining Eq.~\ref{eq:linear_flux_geo} and Eq.~\ref{eq:linear_flux_emb} in the form of linear equations, and fitting parameters using linear regression.

@@

$$
score(w)_{V_{\text{axis}}} = \cos(v_{w}, V_{\text{axis}}) = \frac{u_{w} \cdot V_{\text{axis}}}{\lVert v_{w} \rVert \lVert V_{\text{axis}} \rVert},
$$

where higher scores mean that organization $w$ is more closely aligned to $S^+$ than $S^-$.

>>>>

\begin{equation}
	\label{eq:gravity_basic}
	\hat{T_{ij}} = Cm_{i}m_{j}f(r_{ij}),
\end{equation}
where the \textit{expected flux}, $\hat{T_{ij}}$, or expected number of authors who were affiliated with both organizations during the time period is proportional to the organization's sizes (the sum of both mobile and non-mobile authors, see Data and Methods), $m_{i}$ and $m_{j}$, and decays as a function of their distance, $f(r_{ij})$.
The distance function is defined in terms of either the geographic distance between organizations or the embedding distance. 
For geographic distance, we define $f(r_{ij})$ as the standard power-law function, and for the embedding distance, we use the exponential function, selected as the best performing for each case (Fig.~\ref{fig:supp:predict_emb_bymodel} and Fig.~\ref{fig:supp:predict_geo_bymodel}).
We argue that the performance of a distance measure can be observed from its ability to explain expected mobility between organizations and to make predictions using the gravity model.

@@

\begin{equation}
	\label{eq:gravity_basic}
	\hat{T_{ij}} = Cm_{i}m_{j}f(r_{ij}),
\end{equation}
where $m_{i}$ and $m_{j}$ are the organization masses, or the yearly-average count of unique authors affiliated with each organization during the time window, and $f(r_{ij})$ is a distance-decaying function  between the two organizations. There are two popular forms of the $f(r_{ij})$. The first one is power-law function with $f(r_{ij})= r^{-\alpha}_{ij} \ (\alpha > 0)$, and the another one is exponential function with $f(r_{ij}) = e^{-\beta r_{ij}} \ (\beta > 0)$ \autocite{chen2015distance}.  $f(r_{ij})$, $C$ can be chosen as to fit the data with the simple regression.

We consider separate variants of $f(r_{ij})$ for geographic distance, $g_{ij}$, and the embedding distance, $d_{ij}$, report the best-fit model of each distance.
For geographic distance, we use the power-law function of the gravity law, with $f(g_{ij})= g^{-\alpha}_{ij}$, where $\alpha$ is estimated from the data.
For embedding distance, we use the exponential function, with $f(d_{ij}) = e^{-\beta d_{ij}}$, where $\beta$ is estimated from the data.
To estimate parameters $C$, $\alpha$ and $\beta$ are estimated by log-transforming each variant of \ref{eq:gravity_basic}, obtaining Eq.~\ref{eq:linear_flux_geo} and Eq.~\ref{eq:linear_flux_emb} in the form of linear equations, and fitting parameters using linear regression

@@

$$
score(w)_{V_{\text{axis}}} = \cos(v_{w}, V_{\text{axis}}) = \frac{u_{w} \cdot V_{\text{axis}}}{\lVert v_{w} \rVert \lVert V_{\text{axis}} \rVert},
$$
where higher scores mean that organization $w$ is more closely aligned to $S^+$ than $S^-$.
